On the number of tiles visited by a\\ line segment on a rectangular grid

### 2023-05-22 paper.tex
A few changes before re-submitting to Mathematika

% Authors with different affiliations: from https://tex.stackexchange.com/a/370813/90222
\author[1]{Luis Mendo}
\author[2]{Alex Arkhipov}

\affil[1]{\small Universidad Polit\'ecnica de Madrid. E-mail:  \texttt{luis.mendo@upm.es}}
%Information Processing and Telecommunications Center, Universidad Polit\'ecnica de Madrid\\
%Avenida Complutense, 30. 28040 Madrid, Spain.\\
\affil[2]{\small E-mail: \texttt{arkhipov@alum.mit.edu}}

@@

There is a range of values of $\theta$ for which the above condition is satisfied. This corresponds to the part of the arc with solid line in Figure~\ref{fig: probmax_funt_odd}. For each $\theta$ in this range, the valid positions for the second endpoint of the segment are in a rectangle contained in the tile with lower-left corner $(i-1,j-1)$. These are the positions that cause the first endpoint to be in the reference tile. The rectangle associated with an example $\theta$ is shown in Figure~\ref{fig: probmax_funt_odd}, shaded. The area of this rectangle is the probability $\probmax_{i, j}(\len,\theta)$ that the segment goes from the reference tile to that with lower-left corner $(i-1,j-1)$ conditioned on $\theta$. As seen in the figure, this area is
\begin{equation}
\label{eq: probmax: probmax theta odd}
\probmax_{i, j}(\len, \theta) = \left(\len\cos\theta - i+2\right) \left(\len\sin\theta - j + 2\right).
\end{equation}
The range of allowed values for $\theta$, as can also be deduced from the figure, is $(\theta_0, \theta_1)$ with
\begin{align}
\label{eq: odd theta 0}
\theta_0 & = \arcsin\frac{j-2}{\len}, \\
\label{eq: odd theta 1}
\theta_1 & = \arccos\frac{i-2}{\len}.
\end{align}
For $\theta$ outside of this range the conditional probability $\probmax_{i, j}(\len, \theta)$ is $0$. Averaging $\probmax_{i, j}(\len, \theta)$ over $\theta$ gives the probability that the segment visits this tile, and then multiplying by $4$ yields $\probmax_{i.j}(\len)$:
\begin{equation}
\begin{split}
\label{eq: probmax: cond on theta}
\probmax_{i,j}(\len) &= 4 \int_{\theta_0}^{\theta_1} \frac{\probmax_{i, j}(\len, \theta)}{2\pi} \, \diff \theta \\
&= \frac 2 {\pi} \int_{\arcsin\frac{j-2}{\len}}^{\arccos\frac{i-2}{\len}} \left( \len\cos\theta - i + 2 \right)  \left( \len\sin\theta - j + 2 \right) \, \diff \theta
= g( \len, i-2, j-2).
\end{split}
\end{equation}

@@

Given $\tiles$, the range of lengths $\len$ such that $\tiles = \funt(\len)$ is either \eqref{eq: square, odd max tiles, lengths} or \eqref{eq: square, even max tiles, lengths}. For $\len$ in this range, the number of reachable tiles with lower-left corners $(i-1,j-1)$, $i,j \geq 2$, $i+j-1=\tiles$, i.e.~the number of terms in \eqref{eq: probmax impl}, depends on $\len$. If $\tiles$ is odd the minimum number is $1$, corresponding to a tile with its lower-left corner on the line $x=y$, which gives the first term in the first expression in \eqref{eq: probmax expl}. For $\tiles$ even the minimum number is $2$, corresponding to two symmetrical tiles whose lower-left $x$ and $y$ coordinates differ by $1$, which give the first term in the second expression. Either for $\tiles$ odd or even, additional tiles may become reachable as $\len$ grows (within the range allowed by $\tiles$), and these always occur in symmetrical pairs. These tiles (if any) correspond to the terms in the (possibly empty) sum indexed by $k$ in either expression in \eqref{eq: probmax expl}.

@@

\begin{proposition}
The function $\probmax$ has the following asymptotic behaviour:
\begin{align}
\label{eq: probmax liminf}
\liminf_{\len \rightarrow \infty}\, \probmax(\len) & = 0, \\ 
\label{eq: probmax lim sqrt tiles}
\lim_{\tiles \rightarrow \infty}\, \sqrt{\tiles} \probmax(\len_\tiles) & = \frac {32\sqrt{2}}{105\pi}, \\
\label{eq: probmax limsup sqrt len}
\limsup_{\len \rightarrow \infty}\, \sqrt{\len} \probmax(\len) & = \frac {32\sqrt[4]{2}}{105\pi}.
\end{align}
\end{proposition}

>>>>

% Authors with different affiliations: from https://tex.stackexchange.com/a/370813/90222
\author[1]{Alex Arkhipov}
\author[2]{Luis Mendo}

\affil[1]{\small E-mail: \texttt{arkhipov@alum.mit.edu}}
\affil[2]{\small Universidad Polit\'ecnica de Madrid. E-mail:  \texttt{luis.mendo@upm.es}}
%Information Processing and Telecommunications Center, Universidad Polit\'ecnica de Madrid\\
%Avenida Complutense, 30. 28040 Madrid, Spain.\\

@@

There is a range of values of $\theta$ for which the above condition is satisfied. This corresponds to the part of the arc with solid line in Figure~\ref{fig: probmax_funt_odd}. For each $\theta$ in this range, the valid positions for the second endpoint of the segment are in a rectangle contained in the tile with lower-left corner $(i-1,j-1)$. These are the positions that cause the first endpoint to be in the reference tile. The rectangle associated with an example $\theta$ is shown in Figure~\ref{fig: probmax_funt_odd}, shaded. The area of this rectangle is the probability $\probmax_{i, j}(\len,\theta)$ that the segment goes from the reference tile to that with lower-left corner $(i-1,j-1)$ conditioned on $\theta$. As seen in the figure, this area is
\begin{equation}
\label{eq: probmax: probmax theta odd}
\probmax_{i, j}(\len, \theta) = \left(\len\cos\theta - i+2\right) \left(\len\sin\theta - j + 2\right).
\end{equation}
The range of allowed values for $\theta$, as can also be deduced from the figure, is $(\theta_0, \theta_1)$ with
\begin{align}
\label{eq: odd theta 0}
\theta_0 & = \arcsin\frac{j-2}{\len}, \\
\label{eq: odd theta 1}
\theta_1 & = \arccos\frac{i-2}{\len}.
\end{align}
For $\theta$ outside of this range, the conditional probability $\probmax_{i, j}(\len, \theta)$ is $0$. Averaging $\probmax_{i, j}(\len, \theta)$ over $\theta$ gives the probability that the segment visits this tile, and then multiplying by $4$ yields $\probmax_{i,j}(\len)$:
\begin{equation}
\begin{split}
\label{eq: probmax: cond on theta}
\probmax_{i,j}(\len) &= 4 \int_{\theta_0}^{\theta_1} \frac{\probmax_{i, j}(\len, \theta)}{2\pi} \, \diff \theta \\
&= \frac 2 {\pi} \int_{\arcsin\frac{j-2}{\len}}^{\arccos\frac{i-2}{\len}} \left( \len\cos\theta - i + 2 \right)  \left( \len\sin\theta - j + 2 \right) \, \diff \theta
= g( \len, i-2, j-2).
\end{split}
\end{equation}

@@

Given $\tiles$, the range of lengths $\len$ such that $\tiles = \funt(\len)$ is either \eqref{eq: square, odd max tiles, lengths} or \eqref{eq: square, even max tiles, lengths}. For $\len$ in this range, the number of reachable tiles with lower-left corners $(i-1,j-1)$, $i,j \geq 2$, $i+j-1=\tiles$, i.e.~the number of terms in \eqref{eq: probmax impl}, depends on $\len$. If $\tiles$ is odd, the minimum number is $1$, corresponding to a tile with its lower-left corner on the line $x=y$, which gives the first term in the first expression in \eqref{eq: probmax expl}. For $\tiles$ even, the minimum number is $2$, corresponding to two symmetrical tiles whose lower-left $x$ and $y$ coordinates differ by $1$, which give the first term in the second expression. Either for $\tiles$ odd or even, additional tiles may become reachable as $\len$ grows (within the range allowed by $\tiles$), and these always occur in symmetrical pairs. These tiles (if any) correspond to the terms in the (possibly empty) sum indexed by $k$ in either expression in \eqref{eq: probmax expl}.

@@

\begin{proposition}
The probability $\probmax(\len)$ of visiting the maximum number of tiles on a unit square grid has the following asymptotic behaviour:
\begin{align}
\label{eq: probmax liminf}
\liminf_{\len \rightarrow \infty}\, \probmax(\len) & = 0, \\ 
\label{eq: probmax lim sqrt tiles}
\lim_{\tiles \rightarrow \infty}\, \sqrt{\tiles} \probmax(\len_\tiles) & = \frac {32\sqrt{2}}{105\pi}, \\
\label{eq: probmax limsup sqrt len}
\limsup_{\len \rightarrow \infty}\, \sqrt{\len} \probmax(\len) & = \frac {32\sqrt[4]{2}}{105\pi}.
\end{align}
\end{proposition}
